---
title: "APM1110 - FA 5 - Dacanay"
output: github_document
editor_options: 
  markdown: 
    wrap: 72
---

### 6. An email message can travel through one of three server routes. The percentage of errors in each of the servers and the percentage of messages that travel through each route are shown in the following table. Assume that the servers are independent.
```{r}
library(pander)

Servers <- c("Server 1", "Server 2", "Server 3")
df_email <- data.frame(
  Percentage_of_Messages = c(40, 25, 35),
  Percentage_of_Errors = c(1, 2, 1.5)
)
rownames(df_email) <- Servers

pander(df_email)
```


#### (a) What is the probability of receiving an email containing an error?
We can use the law of total probability. In this problem, we need to get the product of the probability of messages and errors of each server and solve for their sum.
```{r}
# Assigning values for the probability of messages and errors
p_message_1 <- df_email[1, 1] / 100
p_message_2 <- df_email[2, 1] / 100
p_message_3 <- df_email[3, 1] /100

p_error_1 <- df_email[1, 2] / 100
p_error_2 <- df_email[2, 2] / 100
p_error_3 <- df_email[3, 2] / 100

# Product of message and error of each server
prod_server_1 <- p_message_1 * p_error_1
prod_server_2 <- p_message_2 * p_error_2
prod_server_3 <- p_message_3 * p_error_3

# Total sum of all servers
p_email_error <- prod_server_1 + prod_server_2 + prod_server_3
```
##### ANSWER
```{r}
paste0(p_email_error, " or ", p_email_error * 100, "%")
```
The probability of receiving an email containing an error is $0.01425$ or $1.425$%


#### (b) What is the probability that a message will arrive without error?
We can use the complementary approach here, in which we will subtract the probability of receiving an error in email from 1.
```{r}
p_email_wo_error <- 1 - p_email_error
```
##### ANSWER
```{r}
paste0(p_email_wo_error, " or ", p_email_wo_error * 100, "%")
```
The probability of receiving an email containing an error is $0.98575$ or $98.575$%


#### (c) If a message arrives without error, what is the probability that it was sent through server 1?
In this question, we will use the data from (b), which is the probability of receiving an email without an error. We will use Bayes' rule here.
```{r}
p_email_wo_error_1 <- round(((p_message_1 * (1-p_error_1)) / p_email_wo_error), 4)
```
##### ANSWER
```{r}
paste0(p_email_wo_error_1, " or ", p_email_wo_error_1*100, "%")
```
The probability of receiving an email without an error in server 1 is $0.4017$ or $40.17$%



### 9. A software company surveyed managers to determine the probability that they would buy a new graphics package that includes three-dimensional graphics. About 20% of office managers were certain that they would not buy the package, 70% claimed that they would buy, and the others were undecided. Of those who said that they would not buy the package, only 10% said that they were interested in upgrading their computer hardware. Of those interested in buying the graphics package, 40% were also interested in upgrading their computer hardware. Of the undecided, 20% were interested in upgrading their computer hardware. 
```{r}
Decisions <- c("Not buying", "Buying", "Undecided")
df_company <- data.frame(
  Percentage_of_Managers = c(20, 70, 10),
  Percentage_of_Upgrading = c(10, 40, 20)
)
rownames(df_company) <- Decisions

pander(df_company)
```
Let **_A_** denote the intention of not buying, **_B_** the intention of buying, **_C_** the undecided, **_G_** the intention of upgrading the computer hardware, and **_H_** the complementary of **_G_**.


#### (a) Calculate the probability that a manager chosen at random will not upgrade the computer hardware (P(H)).
We can use the complementary approach first of the event that the managers will upgrade the computer hard, then,  the law of total probability. In this problem, we need to get the product of the probability of managers and not upgrading for each decision and solve for their sum.
$$
P(H) = P(A)P(\overline{G}|A) + P(B)P(\overline{G}|B) + P(C)P(\overline{G}|C)
$$
```{r}
# Probability of Manager's decision
p_A <- df_company[1, 1] / 100
p_B <- df_company[2, 1] / 100
p_C <- df_company[3, 1] / 100

# Probability of Upgrading
p_G_A <- df_company[1, 2] / 100
p_G_B <- df_company[2, 2] / 100
p_G_C <- df_company[3, 2] / 100

# Complementary of Upgrading
p_H_A <- 1 - p_G_A
p_H_B <- 1 - p_G_B
p_H_C <- 1 - p_G_C

# Product of the probability of each manager's decision and not upgrading
prod_A_HA <- p_A * p_H_A
prod_B_HB <- p_B * p_H_B
prod_C_HC <- p_C * p_H_C

# Total sum of all managers not upgrading
p_H <- prod_A_HA + prod_B_HB + prod_C_HC
```
##### ANSWER
```{r}
paste0(p_H, " or ", p_H * 100, "%")
```
The probability that a manager chosen at random will not upgrade the computer hardware is $0.68$ or $68$%


#### (b) Explain what is meant by the posterior probability of B given G, P(B|G).
##### EXPLANATION
The posterior probability of **_B_** given **_G_** or **_P(B|G)_** is the probability that **_B_** occurred given that **_G_** occurred. In this problem, it means that if there will be a manager going for upgrading of the computer hardware, what is the chance or probability that that manager was also one of those managers who decided to buy a new graphics package.

To solve this, we'll use some data from (a), that is **_P(A)_**, **_P(B)_**, **_P(C)_**, **_P(G|A)_**, **_P(G|B)_**, and **_P(G|C)_**
$$
P(B|G) = \frac{P(B)P(G|B)}{P(G)}
$$
```{r}
# Product of the probability of each manager's decision and upgrading
prod_A_GA <- p_A * p_G_A
prod_B_GB <- p_B * p_G_B
prod_C_GC <- p_C * p_G_C

# Total sum of all managers going for upgrade
p_G <- prod_A_GA + prod_B_GB + prod_C_GC

# Posterior probability of a manager buying given that he/she is upgrading
p_B_G <- prod_B_GB / p_G
```
##### ANSWER
```{r}
paste0(p_B_G, " or ", p_B_G * 100, "%")
```
The probability of a manager buying a new graphics package given that he/she is upgrading is $0.875$ or $87.5$%.


#### (c)  Construct a tree diagram and use it to calculate the following probabilities: P(G), P(B|G), P(B|H), P(C|G), P(C|H).
```{r}
library(igraph)

vertices <- c("Manager", "P(A)", "P(B)", "P(C)", "P(G|A)", "P(H|A)", "P(G|B)", "P(H|B)", "P(G|C)", "P(H|C)")

edges <- matrix(c(
  "Manager", "P(A)",
  "Manager", "P(B)",
  "Manager", "P(C)",
  "P(A)", "P(G|A)",
  "P(A)", "P(H|A)",
  "P(B)", "P(G|B)",
  "P(B)", "P(H|B)",
  "P(C)", "P(G|C)",
  "P(C)", "P(H|C)"
), byrow = TRUE, ncol = 2)

edge_labels <- c(p_A, p_B, p_C, p_G_A, p_H_A, p_G_B, p_H_B, p_G_C, p_H_C)

g <- graph_from_edgelist(edges, directed = TRUE)
E(g)$label <- edge_labels

V(g)$label <- vertices

layout <- matrix(c(
  0, 0, #manager
  -12, -1, #A
  0, -1, #B
  12, -1, #C
  -20, -2, #G|A
  -12, -2,#H|A
  -4, -2, #G|B
  4, -2, #H|B
  12, -2, #G|C
  20, -2 #H|C
), ncol = 2, byrow = TRUE)

V(g)$x <- layout[,1]
V(g)$y <- layout[,2]

par(mar = c(0, 0, 3, 0))

node_sizes <- c(60,40, 40, 40, 40, 40, 40, 40, 40, 40)

plot(g, layout = layout, edge.label = E(g)$label, edge.label.cex = 1, vertex.label.cex = 0.8,
     vertex.shape = "circle", vertex.size = node_sizes, edge.arrow.size = 0.5,
     main = "A Tree Diagram of Managers' Decisions Probability")

label_coords <- matrix(c(
  -1.1, 0.45,
  1.5, -0.5
), ncol = 2, byrow = TRUE)


text(label_coords[1,1], label_coords[1,2], labels = "Probability of Each Decision", pos = 3, cex = 0.8)
text(label_coords[2,1], label_coords[2,2], labels = "Probability of Upgrading", pos = 3, cex = 0.8)

```

##### CALCULATING P(G)
```{r}
vertices <- c("Manager", "P(A)", "P(B)", "P(C)", "P(G|A)", "P(G|B)", "P(G|C)", "P(G)", (p_A*p_G_A)+(p_B*p_G_B)+(p_C*p_G_C))

edges <- matrix(c(
  "Manager", "P(A)",
  "Manager", "P(B)",
  "Manager", "P(C)",
  "P(A)", "P(G|A)",
  "P(B)", "P(G|B)",
  "P(C)", "P(G|C)",
  "P(G|A)", "P(G)",
  "P(G|B)", "P(G)",
  "P(G|C)", "P(G)",
  "P(G)", (p_A*p_G_A)+(p_B*p_G_B)+(p_C*p_G_C)
), byrow = TRUE, ncol = 2)

edge_labels <- c(p_A, p_B, p_C, p_G_A, p_G_B, p_G_C, p_A*p_G_A, p_B*p_G_B, p_C*p_G_C, "")

g <- graph_from_edgelist(edges, directed = TRUE)
E(g)$label <- edge_labels

V(g)$label <- vertices

layout <- matrix(c(
  0, 0, #manager
  -6, -1, #A
  0, -1, #B
  6, -1, #C
  -6, -2, #G|A
  0, -2, #G|B
  6, -2, #G|C
  0, -3, #G
  5, -3
), ncol = 2, byrow = TRUE)

V(g)$x <- layout[,1]
V(g)$y <- layout[,2]

par(mar = c(0, 0, 3, 0))

node_sizes <- c(60, 40, 40, 40, 40, 40, 40, 40, 40)

plot(g, layout = layout, edge.label = E(g)$label, edge.label.cex = 1, vertex.label.cex = 0.8,
     vertex.shape = "circle", vertex.size = node_sizes, edge.arrow.size = 0.5,
     main = "Total Probability of Upgrading the Computer Hardware")

label_coords <- matrix(c(
  -1.2, 0.65,
  1.7, -0.06,
  -1.1, -0.9
), ncol = 2, byrow = TRUE)


text(label_coords[1,1], label_coords[1,2], labels = "Probability of Each Decision", pos = 3, cex = 0.8)
text(label_coords[2,1], label_coords[2,2], labels = "Probability of Upgrading", pos = 3, cex = 0.8)
text(label_coords[3,1], label_coords[3,2], labels = "Product of each 
decision and upgrading", pos = 3, cex = 0.8)
```


### 13. A malicious spyware can infect a computer system though the Internet or through email. The spyware comes through the Internet 70% of the time and 30% of the time, it gets in through email. If it enters via the Internet the anti-virus detector will detect it with probability 0.6, and via email, it is detected with probability 0.8.
```{r}
Medium <- c("Internet", "Email")
df_spyware <- data.frame(
  Percentage_of_Spyware = c(70, 30),
  Probability_of_Detection = c(0.6, 0.8)
)
rownames(df_spyware) <- Medium

pander(df_spyware)
```
Let **_I_** denote the spyware comes through the Internet while **_E_** the through email. Also, let **_D_** denote the anti-virus detector will detect it while **_N_** the complementary of **_D_**.

#### (a) What is the probability that this spyware infects the system?
Since the probability that the spyware will infect the system is what being asked, it means that we need to get the probability that the anti-virus detector will not detect the spyware. Only when the spyware is not detected will it get to the system and infect it. Hence, we need to get the total probability of the spyware not being detected. That is:
$$
P(N) = P(I) P(N|I) + P(E) P(N|E)
$$
```{r}
# Probability of Spyware Coming Through
p_I <- df_spyware[1, 1] / 100
p_E <- df_spyware[2, 1] / 100

# Probability of Detection
p_D_I <- df_spyware[1, 2]
p_D_E <- df_spyware[2, 2]

# Complementary of Detection
p_N_I <- 1 - p_D_I
p_N_E <- 1 - p_D_E

# Product of the probability of each spyware's medium and its undetectability
prod_I_NI <- p_I * p_N_I
prod_E_NE <- p_E * p_N_E

# Total sum of all managers not upgrading
p_N <- prod_I_NI + prod_E_NE
```
##### ANSWER
```{r}
paste0(p_N, " or ", p_N * 100, "%")
```
The probability that the malicious spyware infects the system is $0.34$ or $34$%.


#### (b) If the spyware is detected, what is the probability that it came through the Internet?
To solve this, we'll use some data from (a), that is **_P(I)_**, **_P(D|I)_**, and **_P(N)_**. Since we have solve for the total probability of the spyware not being detected, we can get its complementary to have the total probability of the spyware being detected.
$$
P(I|D) = \frac{P(I)P(D|I)}{P(\overline{N})}
$$
```{r}
# Complementary of the total probability of the spyware not being detected
p_D <- 1 - p_N

# Computing for the probability that the spyware came through the internet given that it was detected.
p_I_D <- (p_I * p_D_I) / p_D
p_I_D_rounded <- round(p_I_D, 4)
```
##### ANSWER
```{r}
paste0(p_I_D_rounded, " or ", p_I_D_rounded * 100, "%")
```
The probability of a manager buying a new graphics package given that he/she is upgrading is $0.6364$ or $63.64$%.
